---
title: "BISC_2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages}

library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(vegan) #for species diversity calculations

```

```{r load_data}

coral <- read_excel("2018_bisc_coral_cover.xlsx")
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
gps_coord <- read_excel("bisc _rvc_2018 _sites .xlsx")

```




##Is coral cover corelated with percent bleached?

I noticed that there was a lot more bleaching when there was a higher percent coral cover. Now obviously this makes sense off the bat because there can't be much bleaching if there isn't a lot of coral. However it seems like the percentage of bleaching was higher when coral cover was higher. This is exploring that possbility.

##THIS LOOKS AT PERCENT TOTAL BLEACHED OF THE CORAL THAT IS PRESENT###
```{r}

coral <- read_excel("2018_bisc_coral_cover.xlsx")

coral_cover_bleach <- coral %>%
  select(coral_cover, percent_total_bleached) %>%
  filter(coral_cover != 0)

ggplot(coral_cover_bleach, aes(x=coral_cover, y= percent_total_bleached)) +
  geom_point()

coral_cover_bleach_lm <- lm(percent_total_bleached ~ coral_cover, data = coral_cover_bleach)
coral_cover_bleach_lm

#summary of model
summary(coral_cover_bleach_lm)

#data diagnostics
par(mfrow = c(2,2))
plot(coral_cover_bleach_lm)
# qqplot is a little off but not terribly
# resifuals fit is skewed to the right and I'm not quite sure what that means, if it means anything

#had to make percent total bleached into its own vector cause r was freaking out. go figure
percent_total_b <- as.numeric(coral_cover_bleach$percent_total_bleached)

#correlation: person's r
correlation_cover_bleach <- cor.test(coral_cover_bleach$coral_cover, percent_total_b)
correlation_cover_bleach
# not significant. p-value = 0.4654

```

##THIS LOOKS AT PERCENT OF THE NUMBERS OF CORALS BLEACHED. NOT THE AMOUNT BLEACHED##
```{r}


coral_number_bleach <- coral %>%
  select(coral_cover, percent_ind_bleached) %>%
  filter(coral_cover != 0)

ggplot(coral_number_bleach, aes(x=coral_cover, y= percent_ind_bleached)) +
  geom_point()

coral_number_bleach_lm <- lm(percent_ind_bleached ~ coral_cover, data = coral_number_bleach)
coral_number_bleach_lm

#summary of model
summary(coral_number_bleach_lm)

#data diagnostics
par(mfrow = c(2,2))
plot(coral_number_bleach_lm)
# qqplot is a little off but not terribly
# resifuals fit is skewed to the right and I'm not quite sure what that means, if it means anything

#had to make percent total bleached into its own vector cause r was freaking out. go figure
percent_ind_b_vec <- as.numeric(coral_number_bleach$percent_ind_bleached)

#correlation: person's r
correlation_number_bleach <- cor.test(coral_number_bleach$coral_cover, percent_ind_b_vec)
correlation_number_bleach
# not significant. p-value = 0.3009
```






