---
title: "BISC_2018"
author: "Juliette Verstaen"
date: "8/21/2018"
output: html_document
---

```{r load_packages}

library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(vegan) #for species diversity calculations

```

```{r load_data}

TEST_coral <- read_excel("TEST_2018_bisc_coral_health_data.xlsx")
TEST_fish <- read_excel("TEST_RVC_Data_2018 2.xlsx")

TEST_coral_species_numbers <- read_excel("TEST_coral_species_numbers.xlsx")
TEST_coral_cover <- read_excel("TEST_coral_cover.xlsx")


```

##2018 Hogfish Size Distribution 
**last fish data update 08_21_2018**
```{r}


hogfish_test <- TEST_fish %>%
  filter(fish_com == "hogfish") %>%
  select(fish_com,f_1, f_2, f_3, f_4, f_5, f_6, f_7, f_8, f_9) 

hogfish_test[is.na(hogfish_test)] <- 0

hogfish_one <- hogfish_test%>%
  select (f_1) %>%
  filter(f_1 >0)
colnames(hogfish_one) <- c("total")

hogfish_two <- hogfish_test%>%
  select (f_2) %>%
  filter(f_2 >0)
colnames(hogfish_two) <- c("total")

hogfish_three <- hogfish_test%>%
  select (f_3) %>%
  filter(f_3 >0)
colnames(hogfish_three) <- c("total")

hogfish_four <- hogfish_test%>%
  select (f_4) %>%
  filter(f_4 >0)
colnames(hogfish_four) <- c("total")

hogfish_five <- hogfish_test%>%
  select (f_5) %>%
  filter(f_5 >0)
colnames(hogfish_five) <- c("total")

f_1 <- hogfish_one$total
f_2 <- hogfish_two$total
f_3 <- hogfish_three$total
f_4 <- hogfish_four$total
f_5 <- hogfish_five$total

hf_total <- c(f_1, f_2, f_3, f_4, f_5 )
hf_total_vector <- as.numeric(hf_total_vector)

hist(hf_total_vector)

ggplot() +
  aes(hf_total_vector) +
  geom_histogram(binwidth = 1, colour = "navy", fill = "white")+
  xlab("Hogfish Size (cm)")+
  ylab("Count")+
  ggtitle("Hogfish Size Distribution 2018")+
  geom_vline(xintercept = 40.64)+
  theme_bw()


```

##Coral Data Wrangling and Exploration

```{r coral_fun}

coral_bleaching <- TEST_coral %>%
  select("field_number", "quadrat", "coral_species", "percent_live", "percent_bleached") %>%
  filter (coral_species != "NA")

coral_siderastrea <- TEST_coral %>%
  select("field_number", "quadrat", "coral_species", "percent_live", "percent_bleached") %>%
  filter (coral_species != "NA") %>%
  filter(coral_species == "siderastrea_radians" | coral_species == "siderastrea_siderea")

sidsid <- TEST_coral %>%
  select("field_number", "quadrat", "coral_species", "percent_live", "percent_bleached") %>%
  filter (coral_species != "NA") %>%
  filter(coral_species == "siderastrea_siderea")

sid_sid <- sidsid$percent_alive
sid_sid <- as.numeric(sid_sid)


sidrad <- TEST_coral %>%
  select("field_number", "quadrat", "coral_species", "percent_live", "percent_bleached") %>%
  filter (coral_species != "NA") %>%
  filter(coral_species == "siderastrea_radians")

sid_rad <- sidrad$percent_alive
sid_rad <- as.numeric(sid_rad)

###stoped mid thought
graph <- 
  ggplot(data = coral_bleaching, aes(x =coral_species, y = coral_bleaching))+
  geom_bar

graph
  

```



```{r diversity_function}

TEST_coral_species_numbers <- read_excel("TEST_coral_species_numbers.xlsx")

#remember the last line of data here is the total of all sites##
c_shan <- diversity(TEST_coral_species_numbers)
c_shan

#plot(c_shan)

```

##Is coral cover corelated with percent bleached?

I noticed that there was a lot more bleaching when there was a higher percent coral cover. Now obviously this makes sense off the bat because there can't be much bleaching if there isn't a lot of coral. However it seems like the percentage of bleaching was higher when coral cover was higher. This is exploring that possbility.


```{r coral_cover_bleaching}

TEST_coral_cover <- read_excel("TEST_coral_cover.xlsx")

coral_cover_bleach <- TEST_coral_cover %>%
  select(coral_cover, percent_bleached) %>%
  filter(coral_cover != 0)

ggplot(coral_cover_bleach, aes(x=coral_cover, y= percent_bleached)) +
  geom_point()

coral_cover_bleach_lm <- lm(percent_bleached ~ coral_cover, data = coral_cover)
coral_cover_bleach_lm

#summary of model
summary(coral_cover_bleach_lm)

#data diagnostics
par(mfrow = c(2,2))
plot(coral_cover_bleach_lm)
# qqplot is a little off but not terribly
# resifuals fit is skewed to the right and I'm not quite sure what that means, if it means anything


#correlation: person's r
correlation_cover_bleach <- cor.test(coral_cover_bleach$coral_cover, coral_cover_bleach$percent_bleached)
correlation_cover_bleach
# not significant. p-value = 0.2869
```

#Model
Percent Coral Bleaching = 0.08837 -0.66218(Percent Coral Cover_)
R$^2$ = 0.043 
4.3% of variance in percent of coral bleaching is exlained by percent of coral cover. 

#Correlation
There is no significant correlation between coral cover and percent bleached (Pearson's *r* = -0.209???,t(28)= -1.0873, *p* =  0.2869).

##Fish and Coral Correlations

```{r coralcover_fishabundance}

TEST_RVC_Fish_condensed <- read_excel("TEST_RVC_Fish_condensed.xlsx")

fish_coral <- merge(TEST_RVC_Fish_condensed, TEST_coral_cover, by = "site")

#how does coral cover and bleaching effect the hog fish abundance
hog_coral <- fish_coral %>%
  filter( fish_com == "hogfish" | fish_com =="none") %>%
  select(fish_num, f_avg, coral_cover, percent_bleached, percent_diseased, paling) 

hog_coral_lm <- lm(formula = fish_num ~  coral_cover + percent_bleached +percent_diseased + paling, data = hog_coral)
hog_coral_lm

summary(hog_coral_lm)

#how does the coral cover and bleaching effect the game fish abundance (snappers, groupers, hogfish)
gamefish_coral <- fish_coral %>%
  filter( fish_com != "banded_butterflyfish" | fish_com != "foureyed_butterflyfish" | fish_com != "reef_butterflyfish" | fish_com != "spotfin_butterflyfish") %>%
  select(fish_num, f_avg, coral_cover, percent_bleached, percent_diseased, paling) 

gamefish_coral_lm <- lm(formula = fish_num ~ coral_cover + percent_bleached +percent_diseased + paling, data = gamefish_coral)
gamefish_coral_lm

summary(gamefish_coral_lm)

```











