---
title: "BISC_2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages}

library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(vegan) #for species diversity calculations

```

```{r load_data}

coral <- read_excel("2018_bisc_coral_cover.xlsx")
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
gps_coord <- read_excel("bisc _rvc_2018 _sites .xlsx")

```




##Is coral cover corelated with percent bleached?

I noticed that there was a lot more bleaching when there was a higher percent coral cover. Now obviously this makes sense off the bat because there can't be much bleaching if there isn't a lot of coral. However it seems like the percentage of bleaching was higher when coral cover was higher. This is exploring that possbility.

##THIS LOOKS AT PERCENT TOTAL BLEACHED OF THE CORAL THAT IS PRESENT###
```{r}

coral <- read_excel("2018_bisc_coral_cover.xlsx")

coral_cover_bleach <- coral %>%
  select(coral_cover, percent_total_bleached) %>%
  filter(coral_cover != 0)

ggplot(coral_cover_bleach, aes(x=coral_cover, y= percent_total_bleached)) +
  geom_point()

coral_cover_bleach_lm <- lm(percent_total_bleached ~ coral_cover, data = coral_cover_bleach)
coral_cover_bleach_lm

#summary of model
summary(coral_cover_bleach_lm)

#data diagnostics
par(mfrow = c(2,2))
plot(coral_cover_bleach_lm)
# qqplot is a little off but not terribly
# resifuals fit is skewed to the right and I'm not quite sure what that means, if it means anything

#had to make percent total bleached into its own vector cause r was freaking out. go figure
percent_total_b <- as.numeric(coral_cover_bleach$percent_total_bleached)

#correlation: person's r
correlation_cover_bleach <- cor.test(coral_cover_bleach$coral_cover, percent_total_b)
correlation_cover_bleach
# not significant. p-value = 0.4654

```

##THIS LOOKS AT PERCENT OF THE NUMBERS OF CORALS BLEACHED. NOT THE AMOUNT BLEACHED##
```{r}

coral_number_bleach <- coral %>%
  select(coral_cover, percent_ind_bleached) %>%
  filter(coral_cover != 0)

ggplot(coral_number_bleach, aes(x=coral_cover, y= percent_ind_bleached)) +
  geom_point()

coral_number_bleach_lm <- lm(percent_ind_bleached ~ coral_cover, data = coral_number_bleach)
coral_number_bleach_lm

#summary of model
summary(coral_number_bleach_lm)

#data diagnostics
par(mfrow = c(2,2))
plot(coral_number_bleach_lm)
# qqplot is a little off but not terribly
# resifuals fit is skewed to the right and I'm not quite sure what that means, if it means anything

#had to make percent total bleached into its own vector cause r was freaking out. go figure
percent_ind_b_vec <- as.numeric(coral_number_bleach$percent_ind_bleached)

#correlation: person's r
correlation_number_bleach <- cor.test(coral_number_bleach$coral_cover, percent_ind_b_vec)
correlation_number_bleach
# not significant. p-value = 0.3009
```
##THIS LOOKS AT PERCENT OF THE NUMBERS OF CORALS BLEACHED AND/OR PALING. NOT THE AMOUNT BLEACHED##

```{r}

coral_bleach_pale <- coral %>%
  select(coral_cover, percent_p_b) %>%
  filter(coral_cover != 0)

ggplot(coral_bleach_pale, aes(x=coral_cover, y= percent_p_b)) +
  geom_point()

coral_bleach_pale_lm <- lm(percent_p_b ~ coral_cover, data = coral_bleach_pale)
coral_bleach_pale_lm

#summary of model
summary(coral_bleach_pale_lm)

#data diagnostics
par(mfrow = c(2,2))
plot(coral_bleach_pale_lm)
# qqplot is a little off but not terribly
# resifuals fit is skewed to the right and I'm not quite sure what that means, if it means anything

#had to make percent total bleached into its own vector cause r was freaking out. go figure
percent_p_b_vec <- as.numeric(coral_bleach_pale$percent_p_b)

#correlation: person's r
correlation_bleach_pale <- cor.test(coral_bleach_pale$coral_cover, percent_p_b_vec)
correlation_bleach_pale
# not significant. p-value = 0.2712
```


  ##IS CORAL COVER CORRELATED WITH GROUPER, HOGFISH, and SNAPPER NUMBERS##
  
```{r}

###########################################################################
####this is the same chunk of code for all coral and game fish analysis###
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
coral <- read_excel("2018_bisc_coral_cover.xlsx")

fish_coral <- merge(fish, coral, by = "site")

#how does the coral cover and bleaching effect the game fish abundance (snappers, groupers, hogfish)
gamefish_coral <- fish_coral %>%
  filter( fish_com != "banded_butterflyfish" | fish_com != "foureyed_butterflyfish" | fish_com != "reef_butterflyfish" | fish_com != "spotfin_butterflyfish" | fish_com != "none") %>%
  select(fish_num, f_avg, coral_cover, percent_total_bleached, percent_ind_bleached, percent_paling, paling_coral, percent_p_b) 
##########################################################################

#correlation: person's r
correlation_game_coralcover <- cor.test(gamefish_coral$coral_cover, gamefish_coral$fish_num)
correlation_game_coralcover
# not significant. p-value = 0.9246
#cor = 0.0079297

```
  
## IS THERE A CORRELATIONS BETWEEN PERCENT OF CORALS THAT ARE PALING AND/OR BLEACHING AND GAME FISH
```{r}

###########################################################################
####this is the same chunk of code for all coral and game fish analysis###
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
coral <- read_excel("2018_bisc_coral_cover.xlsx")

fish_coral <- merge(fish, coral, by = "site")

#how does the coral cover and bleaching effect the game fish abundance (snappers, groupers, hogfish)
gamefish_coral <- fish_coral %>%
  filter( fish_com != "banded_butterflyfish" | fish_com != "foureyed_butterflyfish" | fish_com != "reef_butterflyfish" | fish_com != "spotfin_butterflyfish" | fish_com != "none") %>%
  select(fish_num, f_avg, coral_cover, percent_total_bleached, percent_ind_bleached, percent_paling, paling_coral, percent_p_b) 
##########################################################################

percent_p_b_vec_2 <- as.numeric(gamefish_coral$percent_p_b)

#correlation: person's r
correlation_game_percentpb <- cor.test(gamefish_coral$coral_cover, percent_p_b_vec_2)
correlation_game_percentpb
# significant. p-value = 0.00461
#cor = -0.2498671. very weak/bacisally no correlation
```

  
## DOES CORAL HEALTH (cover, % of coral bleaching, percent of coral diseased, whether or not paling) EFFECT GAME FISH NUMBERS##

```{r}

###########################################################################
####this is the same chunk of code for all coral and game fish analysis###
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
coral <- read_excel("2018_bisc_coral_cover.xlsx")

fish_coral <- merge(fish, coral, by = "site")

#how does the coral health effect the game fish abundance (snappers, groupers, hogfish)
gamefish_coral <- fish_coral %>%
  filter( fish_com != "banded_butterflyfish" | fish_com != "foureyed_butterflyfish" | fish_com != "reef_butterflyfish" | fish_com != "spotfin_butterflyfish" | fish_com != "none") %>%
  select(fish_num, f_avg, coral_cover, percent_total_bleached, percent_ind_bleached, paling_coral, percent_paling, percent_p_b) 
##########################################################################

gamefish_coral_lm_percentb_binarypale <- lm(formula = fish_num ~ coral_cover + percent_total_bleached + paling_coral, data = gamefish_coral)
gamefish_coral_lm_percentb_binarypale

summary(gamefish_coral_lm_percentb_binarypale)


##REMOVED NAS FORM CORAL BLEACHING# means that removing the data where there is no coral cover, I don't love that/ will try to find a way around it
#how does the coral health effect the game fish abundance (snappers, groupers, hogfish)
gamefish_coral_removeNAlines <- fish_coral %>%
  filter( fish_com != "banded_butterflyfish" | fish_com != "foureyed_butterflyfish" | fish_com != "reef_butterflyfish" | fish_com != "spotfin_butterflyfish" | fish_com != "none") %>%
  select(fish_num, f_avg, coral_cover, percent_total_bleached, percent_ind_bleached, paling_coral, percent_paling, percent_p_b) %>%
  filter(percent_total_bleached != "NA")
##########################################################################

gamefish_coral_lm_percentb_binarypale_noNA <- lm(formula = fish_num ~ coral_cover + percent_total_bleached + paling_coral, data = gamefish_coral)
gamefish_coral_lm_percentb_binarypale_noNA

summary(gamefish_coral_lm_percentb_binarypale_noNA)
```

## DOES CORAL HEALTH (cover, percent of coral experiencing bleaching, disease, paling) EFFECT GAME FISH NUMBERS##

  
  

