---
title: "BISC_2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages}

library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(vegan) #species diversity calculations
library(sf) #spatial
library(sp) #spatial
library(gstat) #spatial
library(rgdal) #spatial
library(factoextra) #cluster
library(cluster) #cluster
library(magrittr) #cluster


```

```{r load_data}

coral <- read_excel("2018_bisc_coral_cover.xlsx")
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
gps_coord <- read_excel("bisc _rvc_2018 _sites .xlsx")

```


##Species Richness Test###

```{r}
coral_species_numbers <- read_excel("2018_bisc_coral_species_numbers.xlsx")

#remember the last line of data here is the total of all sites##
c_shan <- diversity(coral_species_numbers)
c_shan

plot(c_shan)
```


##Is coral cover corelated with percent bleached?

I noticed that there was a lot more bleaching when there was a higher percent coral cover. Now obviously this makes sense off the bat because there can't be much bleaching if there isn't a lot of coral. However it seems like the percentage of bleaching was higher when coral cover was higher. This is exploring that possbility.

##THIS LOOKS AT PERCENT TOTAL BLEACHED OF THE CORAL THAT IS PRESENT###
```{r}

coral <- read_excel("2018_bisc_coral_cover.xlsx")

coral_cover_bleach <- coral %>%
  select(coral_cover, percent_total_bleached) %>%
  filter(coral_cover != 0)

ggplot(coral_cover_bleach, aes(x=coral_cover, y= percent_total_bleached)) +
  geom_point()

coral_cover_bleach_lm <- lm(percent_total_bleached ~ coral_cover, data = coral_cover_bleach)
coral_cover_bleach_lm

#summary of model
summary(coral_cover_bleach_lm)

#data diagnostics
par(mfrow = c(2,2))
plot(coral_cover_bleach_lm)
# qqplot is a little off but not terribly
# resifuals fit is skewed to the right and I'm not quite sure what that means, if it means anything

#had to make percent total bleached into its own vector cause r was freaking out. go figure
percent_total_b <- as.numeric(coral_cover_bleach$percent_total_bleached)

#correlation: person's r
correlation_cover_bleach <- cor.test(coral_cover_bleach$coral_cover, percent_total_b)
correlation_cover_bleach
# not significant. p-value = 0.2397

```

####how did i do speciec richness and coral cover correlation... need to remember
```{r}

#is coral cover correltated with species richness
correlation_cover_richness <- cor.test(coral$coral_cover, c_shan)
correlation_cover_richness
#not significantly correlated p - 0.4029187

#is paling/bleaching correlated to species richness

correlation_cover_richness <- cor.test(coral_cover_bleach$percent_p_b, c_shan)
correlation_cover_richness

```





##THIS LOOKS AT PERCENT OF THE NUMBERS OF CORALS BLEACHED. NOT THE AMOUNT BLEACHED##
```{r}

coral_number_bleach <- coral %>%
  select(coral_cover, percent_ind_bleached) %>%
  filter(coral_cover != 0)

ggplot(coral_number_bleach, aes(x=coral_cover, y= percent_ind_bleached)) +
  geom_point()

coral_number_bleach_lm <- lm(percent_ind_bleached ~ coral_cover, data = coral_number_bleach)
coral_number_bleach_lm

#summary of model
summary(coral_number_bleach_lm)

#data diagnostics
par(mfrow = c(2,2))
plot(coral_number_bleach_lm)
# qqplot is a little off but not terribly
# resifuals fit is skewed to the right and I'm not quite sure what that means, if it means anything

#had to make percent total bleached into its own vector cause r was freaking out. go figure
percent_ind_b_vec <- as.numeric(coral_number_bleach$percent_ind_bleached)

#correlation: person's r
correlation_number_bleach <- cor.test(coral_number_bleach$coral_cover, percent_ind_b_vec)
correlation_number_bleach
# not significant. p-value = 0.3009
```


##THIS LOOKS AT PERCENT OF THE NUMBERS OF CORALS BLEACHED AND/OR PALING. NOT THE AMOUNT BLEACHED##

```{r}

coral_bleach_pale <- coral %>%
  select(coral_cover, percent_p_b) %>%
  filter(coral_cover != 0)

ggplot(coral_bleach_pale, aes(x=coral_cover, y= percent_p_b)) +
  geom_point()

coral_bleach_pale_lm <- lm(percent_p_b ~ coral_cover, data = coral_bleach_pale)
coral_bleach_pale_lm

#summary of model
summary(coral_bleach_pale_lm)

#data diagnostics
par(mfrow = c(2,2))
plot(coral_bleach_pale_lm)
# qqplot is a little off but not terribly
# resifuals fit is skewed to the right and I'm not quite sure what that means, if it means anything

#had to make percent total bleached into its own vector cause r was freaking out. go figure
percent_p_b_vec <- as.numeric(coral_bleach_pale$percent_p_b)

#correlation: person's r
correlation_bleach_pale <- cor.test(coral_bleach_pale$coral_cover, percent_p_b_vec)
correlation_bleach_pale
# not significant. p-value = 0.2712
```


##IS CORAL COVER CORRELATED WITH GAMEFISH (GROUPER, HOGFISH, and SNAPPER NUMBERS)
  
```{r}

###########################################################################
####this is the same chunk of code for all coral and game fish analysis###
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
coral <- read_excel("2018_bisc_coral_cover.xlsx")

fish_coral <- merge(fish, coral, by = "site")

#how does the coral cover and bleaching effect the game fish abundance (snappers, groupers, hogfish)
gamefish_coral <- fish_coral %>%
  filter( fish_com != "banded_butterflyfish" | fish_com != "foureyed_butterflyfish" | fish_com != "reef_butterflyfish" | fish_com != "spotfin_butterflyfish" | fish_com != "none") %>%
  select(fish_num, f_avg, coral_cover, percent_total_bleached, percent_ind_bleached, percent_paling, paling_coral, percent_p_b) 
##########################################################################

#correlation: person's r
correlation_game_coralcover <- cor.test(gamefish_coral$coral_cover, gamefish_coral$fish_num)
correlation_game_coralcover
# not significant. p-value = 0.9246
#cor = 0.0079297

```
  
## IS THERE A CORRELATIONS BETWEEN PERCENT OF CORALS THAT ARE PALING AND/OR BLEACHING AND GAME FISH
```{r}

###########################################################################
####this is the same chunk of code for all coral and game fish analysis###
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
coral <- read_excel("2018_bisc_coral_cover.xlsx")

fish_coral <- merge(fish, coral, by = "site")

#how does the coral cover and bleaching effect the game fish abundance (snappers, groupers, hogfish)
gamefish_coral <- fish_coral %>%
  filter( fish_com != "banded_butterflyfish" | fish_com != "foureyed_butterflyfish" | fish_com != "reef_butterflyfish" | fish_com != "spotfin_butterflyfish" | fish_com != "none") %>%
  select(fish_num, f_avg, coral_cover, percent_total_bleached, percent_ind_bleached, percent_paling, paling_coral, percent_p_b) 
##########################################################################

percent_p_b_vec_2 <- as.numeric(gamefish_coral$percent_p_b)

#correlation: person's r
correlation_game_percentpb <- cor.test(gamefish_coral$coral_cover, percent_p_b_vec_2)
correlation_game_percentpb
# significant. p-value = 0.00461
#cor = -0.2498671. very weak/bacisally no correlation
```

  
## DOES CORAL HEALTH (cover, % of coral bleaching, percent of coral diseased, whether or not paling) EFFECT GAME FISH NUMBERS##

### took two approaches, 1: removing NAs, 2: replacing them with -1
```{r}

##########################################################################
###########################################################################
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
coral_negs <- read_excel("2018_bisc_coral_cover_negs.xlsx")

fish_coral <- merge(fish, coral_negs, by = "site")

#how does the coral health effect the game fish abundance (snappers, groupers, hogfish)
gamefish_coral_negs <- fish_coral %>%
  filter( fish_com != "banded_butterflyfish" | fish_com != "foureyed_butterflyfish" | fish_com != "reef_butterflyfish" | fish_com != "spotfin_butterflyfish" | fish_com != "none") %>%
  select(fish_num, f_avg, coral_cover, percent_total_bleached, percent_ind_bleached, paling_coral, percent_paling, percent_p_b) 
##########################################################################

gamefish_coral_lm_percentb_binarypale <- lm(formula = fish_num ~ coral_cover + percent_total_bleached + paling_coral, data = gamefish_coral)
gamefish_coral_lm_percentb_binarypale

summary(gamefish_coral_lm_percentb_binarypale)

##########################################################################
##########################################################################
##REMOVED NAS FORM CORAL BLEACHING# means that removing the data where there is no coral cover, I don't love that/ will try to find a way around it
#how does the coral health effect the game fish abundance (snappers, groupers, hogfish)
gamefish_coral_removeNAlines <- fish_coral %>%
  filter( fish_com != "banded_butterflyfish" | fish_com != "foureyed_butterflyfish" | fish_com != "reef_butterflyfish" | fish_com != "spotfin_butterflyfish" | fish_com != "none") %>%
  select(fish_num, f_avg, coral_cover, percent_total_bleached, percent_ind_bleached, paling_coral, percent_paling, percent_p_b) %>%
  filter(percent_total_bleached != "NA")
##########################################################################

gamefish_coral_lm_percentb_binarypale_noNA <- lm(formula = fish_num ~ coral_cover + percent_total_bleached + paling_coral, data = gamefish_coral)
gamefish_coral_lm_percentb_binarypale_noNA

summary(gamefish_coral_lm_percentb_binarypale_noNA)
```

## DOES CORAL HEALTH (cover, percent of coral experiencing bleaching, disease, paling) EFFECT GAME FISH NUMBERS##
- going to continue on with -1 instead of removing NAs

```{r}

##########################################################################
###########################################################################
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
coral_negs <- read_excel("2018_bisc_coral_cover_negs.xlsx")

fish_coral <- merge(fish, coral_negs, by = "site")

#how does the coral health effect the game fish abundance (snappers, groupers, hogfish)
gamefish_coral_negs <- fish_coral %>%
  filter( fish_com != "banded_butterflyfish" | fish_com != "foureyed_butterflyfish" | fish_com != "reef_butterflyfish" | fish_com != "spotfin_butterflyfish" | fish_com != "none") %>%
  select(fish_num, f_avg, coral_cover, percent_total_bleached, percent_ind_bleached, paling_coral, percent_paling, percent_p_b) 
##########################################################################

gamefish_coral_lm_percentindbleachpale <- lm(formula = fish_num ~ coral_cover + percent_ind_bleached + percent_paling, data = gamefish_coral)
gamefish_coral_lm_percentindbleachpale

summary(gamefish_coral_lm_percentindbleachpale)
```


##PCA (Attempts)

```{r}

gamefish_coral_PCA <- prcomp(gamefish_coral_negs, scale=TRUE)
gamefish_coral_PCA

plot(gamefish_coral_PCA)

summary(gamefish_coral_PCA)

biplot(gamefish_coral_PCA)

```

##Cluster Analysis
- need to relearn how to import maps and lay the data over it 
- i will try to show a color range indicating the amount of bleaching/paling 
- don't know how i can incorporate fish into this...

```{r}

##########################################################################
###########################################################################
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
coral_negs <- read_excel("2018_bisc_coral_cover_negs.xlsx")

fish_coral <- merge(fish, coral_negs, by = "site")

#how does the coral health effect the game fish abundance (snappers, groupers, hogfish)
gamefish_coral_negs <- fish_coral %>%
  filter( fish_com != "banded_butterflyfish" | fish_com != "foureyed_butterflyfish" | fish_com != "reef_butterflyfish" | fish_com != "spotfin_butterflyfish" | fish_com != "none") %>%
  select(fish_num, f_avg, coral_cover, percent_total_bleached, percent_ind_bleached, paling_coral, percent_paling, percent_p_b) 
##########################################################################

###don't even know what im doing here####
scale(gamefish_coral_negs)
res.dist <- get_dist(gamefish_coral_negs, stand = TRUE, method = "pearson")
fviz_dist(res.dist, 
   gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))

fviz_nbclust(gamefish_coral_negs, kmeans, method = "gap_stat")

#http://www.sthda.com/english/articles/25-cluster-analysis-in-r-practical-guide/111-types-of-clustering-methods-overview-and-quick-start-r-code/
##handy code for clustering

```





##Spatial Mapping and Analysis
get back to this, whole mess of not working

```{r}
################################################################
######Combine attribute and lat/lon data#######################
################################################################
gps_coord <- read_excel("bisc _rvc_2018 _sites .xlsx")
fish <- read_excel("2018_bisc_RVC_Fish_condensed.xlsx")
coral_negs <- read_excel("2018_bisc_coral_cover_negs.xlsx")

fish_coral <- merge(fish, coral_negs, by = "site")
gps_fish_coral <- merge(fish_coral, gps_coord, by = "site")

bisc_ben <- st_read(dsn = ".", layer = "biscben")
parks <- st_read(dsn = ".", layer = "nps_boundary")
states <- st_read(dsn = ".", layer = "cb_2017_us_state_20m")

df_fishcoral <- st_as_sf(gps_fish_coral, coords = c("lat","long"))

st_crs(bisc_ben) <- 4326
st_crs(parks) <- 4326
st_crs(states) <- 4326

st_transform(parks, 4326)
st_transform(bisc_ben, 4326)
st_transform(states, 4326)

################################
####exploratory ploting########
###############################

percent_p_b_spatial <- df_fishcoral  %>% 
  dplyr::select(percent_p_b)

plot(percent_p_b_spatial) 
plot(states) 

############################
#########ggploting##########
############################
ggplot(percent_p_b_spatial) +
  geom_sf(data = bisc_ben, fill = "gray60") +
  geom_sf(aes(color = percent_p_b_spatial)) +
  scale_color_gradient(low = "yellow", high = "red") +
  theme_bw() +
  ggtitle("California PM2.5 (12/10/2017)") +
  theme(legend.position = "none") # Woohoo!
```



## Kriging
- don't think i will do this because I'm not asking a predictive questions ex: "based on bleaching at site x and y what is bleaching like at site z"





  
  

